---
title: "Great lakes_Datahack"
author: "Saurabh Khodake"
date: "January 13, 2018"
output: html_document
---

##Creating training and testing datasets  


```{r, include=FALSE}
setwd("C:\\Users\\nirva\\IdeaProjects\\Analytics Vidhya/Great lakes sale")
pp_train<-read.csv("train.csv")
pp_test<-read.csv("test.csv")
pp_temp1<-read.csv("test.csv")
head(pp_train,n=1)
head(pp_test,n=1)
pp_test$SALES_PRICE<-NA
pp<-rbind(pp_train,pp_test)
pp<-pp[,-1]
```

##Understanding the variables

```{r}
str(pp)
summary(pp)

```
##Functions for histograms
```{r, include=FALSE}
hplt=function(data,x)
{
for(i in x)
{
hp<-ggplot(data,aes(x=data[,x]))+geom_histogram(bins = 30,na.rm = T)+
xlab(colnames(data[x]))+ylab("Count")
return(hp)
}
}

dhplots <- function(data,fun,ncol) 
{
col=ncol(data)
k=1
j<-numeric()
for(i in 1:col)
{
if(is.numeric(data[,i])| (is.integer(data[,i]))==TRUE)
  {
j[k]<-i
k=k+1
  }
}
plt <- list()
for (i in j) 
  {
  p <- fun(data=data,x=i)
  plt<- c(plt, list(p))
}
m1<-marrangeGrob(plt, ncol=ncol,nrow=2)
m1
}
```

##Functions for density plots

```{r}
dplt=function(data,x)
{
for(i in x)
{
dp<-ggplot(data,aes(x=data[,x]))+geom_density(na.rm = T)+
xlab(colnames(data[x]))+ylab("Count")
return(dp)
} 
}

ddplots <- function(data,fun,ncol) 
{
col=ncol(data)
k=1
j<-numeric()
for(i in 1:col)
{
if(is.numeric(data[,i])| (is.integer(data[,i]))==TRUE)
{
j[k]<-i
k=k+1
}
}
plt <- list()
  for (i in j) {
    p <- fun(data=data,x=i)
    plt<- c(plt, list(p))
  }
m1<-marrangeGrob(plt, ncol=ncol,nrow=2)
m1
}

```


##Histogram
```{r}
require(ggplot2)
require(gridExtra)
dhplots(fun=hplt,data = pp_train[,-1],ncol = 2)
dhplots(fun=hplt,data = pp_test[,-1],ncol = 2)
```

##Density plots
```{r, eval=FALSE, include=FALSE}
ddplots(fun=dplt,data=pp_train[,-1],ncol = 2)
ddplots(fun=dplt,data=pp_test[,-1],ncol = 2)

```


##Area variable

```{r, warning=FALSE}
library(dplyr)
library(sqldf)
library(ggplot2)
library(gridExtra)
unique(pp$AREA)
pp$AREA<-as.character(pp$AREA)
pp<-pp %>% mutate(AREA=ifelse(AREA=="Ann Nagar" | AREA=="Ana Nagar",yes = "Anna Nagar",no = AREA))
pp<-pp %>% mutate(AREA=ifelse(AREA=="Chrmpet" | AREA=="Chormpet" | AREA=="Chrompt",yes ="Chrompet",no =AREA))
pp<-pp %>% mutate(AREA=ifelse(AREA=="TNagar",yes = "T Nagar",no = AREA))
pp<-pp %>% mutate(AREA=ifelse(AREA=="Karapakam",yes ="Karapakkam",no = AREA))
pp<-pp %>% mutate(AREA=ifelse(AREA=="Adyr",yes = "Adyar",no = AREA))
pp<-pp %>% mutate(AREA=ifelse(AREA=="Velchery",yes = "Velachery",no = AREA))
pp<-pp %>% mutate(AREA=ifelse(AREA=="KKNagar",yes = "KK Nagar",no = AREA))
unique(pp$AREA)
pp$AREA<-as.factor(pp$AREA)
SPAR<-sqldf("select AREA,avg(SALES_PRICE) as SP  from pp group by 1 ")
SPAR$SP
SPAR_br<-ggplot(data=SPAR,aes(x=AREA,y=SP/1000))+geom_col(fill="blue",position = "dodge",width = 0.5)
SPAR_bx<-ggplot(data=pp,aes(x=AREA,y=SALES_PRICE/1000))+geom_boxplot(fill="red",position = "dodge")
grob2<-list(SPAR_br,SPAR_bx)
marrangeGrob(grob2,nrow = 2,ncol = 1)
```

##Interior sq ft variable (INT_SQFT)

```{r, message=FALSE, warning=FALSE}
summary(pp$INT_SQFT)
SPIS<-sqldf("select INT_SQFT,avg(SALES_PRICE) as SP  from pp group by 1 ")
SPIS_hg<-ggplot(data=SPIS,aes(x=INT_SQFT))+geom_histogram(fill="blue",binwidth = 10)
SPIS_sc<-ggplot(data=pp,aes(x=INT_SQFT,y=SALES_PRICE/1000))+geom_point(color="red")+geom_smooth(method=lm,se=FALSE)
grob2<-list(SPIS_hg,SPIS_sc)
marrangeGrob(grob2,nrow = 2,ncol = 1)

```

##Plot of Area,Int_sq.ft and sales price

```{r, warning=FALSE}
SPARIS_bx<-ggplot(data=pp,aes(x=AREA,y=SALES_PRICE/1000,col=INT_SQFT))+geom_col()
SPARIS_bx
SPARIS_ji<-ggplot(data=pp,aes(x=INT_SQFT,y=SALES_PRICE/1000,col=AREA))+geom_jitter(size=1)
grob3<-list(SPARIS_bx,SPARIS_ji)
marrangeGrob(grobs = grob3,nrow=2,ncol =1 )
```


##DATE SALE variable

```{r, include=FALSE}
library(lubridate)
pp$DATE_SALE<-dmy(pp$DATE_SALE)

```

##Date Build variable

```{r}
pp$DATE_BUILD<-dmy(pp$DATE_BUILD)

```

##Creating a new variable---No of years before sale

```{r}
pp$total_yrs<-as.numeric(difftime(pp$DATE_SALE,pp$DATE_BUILD,units = "days"))
summary(pp$total_yrs)
par(mfrow=c(1,2))
hist(pp$total_yrs,col="red")
hist(sqrt(pp$total_yrs),col = "blue")

```

##Distance from the mainroad variable---DIST_MAINROAD

```{r, include=FALSE}
summary(pp$DIST_MAINROAD)
hist(pp$DIST_MAINROAD,breaks =100,col = "blue")
```

##Number of bedrooms_N_BEDROOM

```{r}
summary(pp$N_BEDROOM)
pp[is.na(pp$N_BEDROOM),]
hist(pp$N_BEDROOM,col = "blue")
table(pp$N_BEDROOM)
```

##Number of bathrooms---N_BATHROOM

```{r}
summary(pp$N_BATHROOM)
pp[is.na(pp$N_BATHROOM),]
hist(pp$N_BATHROOM,col = "blue")
table(pp$N_BATHROOM)
```

##Total Number of Rooms--N_ROOM

```{r}
summary(pp$N_ROOM)
hist(pp$N_ROOM,col = "blue")
table(pp$N_ROOM)

```

##Imputing missing values in N_BEDROOM and N_BATHROOM

```{r}
library(ggplot2)

ggplot(data = pp,aes(x =N_BEDROOM,fill=as.factor(N_BATHROOM)))+geom_bar()

ggplot(data = pp,aes(x =N_ROOM,fill=as.factor(N_BEDROOM)))+geom_bar(width = 0.5)

##So if the bedroom is 1 or 2 bathroom is mostly 1

pp[is.na(pp$N_BATHROOM),"N_BATHROOM"]<-1

##SO if the total number of rooms is 4 most of the houses have 2 Bedrooms

pp[is.na(pp$N_BEDROOM),"N_BEDROOM"]<-2

```

##Sale Condtion variable_SALE_COND

```{r}
pp$SALE_COND<-as.character(pp$SALE_COND)
unique(pp$SALE_COND)
pp<-pp %>% mutate(SALE_COND=ifelse(SALE_COND=="Partiall" | SALE_COND=="PartiaLl",yes = "Partial",no =SALE_COND))
pp<-pp %>% mutate(SALE_COND=ifelse(SALE_COND=="Ab Normal",yes = "AbNormal",no =SALE_COND))
pp<-pp %>% mutate(SALE_COND=ifelse(SALE_COND=="Adj Land",yes = "AdjLand",no =SALE_COND))
unique(pp$SALE_COND)
pp$SALE_COND<-as.factor(pp$SALE_COND)

```

##Park facilities variable_PARK_FACIL

```{r}
levels(pp$PARK_FACIL)
levels(pp$PARK_FACIL)[2]<-levels(pp$PARK_FACIL)[1]
table(pp$PARK_FACIL)
SPPF<-sqldf("select PARK_FACIL,avg(SALES_PRICE) as SP  from pp group by PARK_FACIL ")
ggplot(data=SPPF,aes(x=PARK_FACIL,y=SP/1000))+geom_bar(fill="darkblue",stat = "Identity",width = 0.5)
```

##Building type variable_BUILDTYPE

```{r, warning=FALSE}
levels(pp$BUILDTYPE)
pp$BUILDTYPE<-as.character(pp$BUILDTYPE)
unique(pp$BUILDTYPE)
pp<-pp %>% mutate(BUILDTYPE=ifelse(BUILDTYPE=="Commercil" | BUILDTYPE=="Comercial",yes="Commercial",no=BUILDTYPE))
pp<-pp %>% mutate(BUILDTYPE=ifelse(BUILDTYPE=="Other",yes="Others",no=BUILDTYPE))
unique(pp$BUILDTYPE)
pp$BUILDTYPE<-as.factor(pp$BUILDTYPE)

##PLOTS

SPBT<-sqldf("select BUILDTYPE,avg(SALES_PRICE) as SP  from pp group by BUILDTYPE ")
SPBT<-ggplot(data=pp,aes(x=BUILDTYPE,y=SALES_PRICE/1000))+geom_bar(width=0.5,stat="Identity",color="darkblue")
SPISBT<-ggplot(data=pp,aes(x=INT_SQFT,y=SALES_PRICE/1000,col=BUILDTYPE))+geom_jitter(size=1)
grob2<-list(SPBT,SPISBT)
marrangeGrob(grob2,nrow = 1,ncol =1)

```

##UTILITY_AVAIL

```{r, message=FALSE, warning=FALSE}
levels(pp$UTILITY_AVAIL)
levels(pp$UTILITY_AVAIL)[1]<-levels(pp$UTILITY_AVAIL)[2]
SPUA<-sqldf("select UTILITY_AVAIL,avg(SALES_PRICE) as SP from pp group by UTILITY_AVAIL ")
SPUA_br<-ggplot(data=SPUA,aes(x=UTILITY_AVAIL,y=SP/1000))+geom_bar(width=0.5,stat="Identity",fill="darkblue")
SPUABT<-sqldf("select UTILITY_AVAIL,avg(SALES_PRICE) as SP,BUILDTYPE from pp group by UTILITY_AVAIL,BUILDTYPE ")
SPUABT_br<-ggplot(data=pp,aes(x=UTILITY_AVAIL,y=SALES_PRICE/1000,fill=BUILDTYPE))+geom_bar(width=0.5,stat="Identity")
grob4<-list(SPUA_br,SPUABT_br)
marrangeGrob(grob4,nrow = 2,ncol =1)

```

##STREET variable

```{r}
pp$STREET<-as.character(pp$STREET)
unique(pp$STREET)
pp<-pp %>% mutate(STREET=ifelse(STREET=="No Access",yes="NoAccess",no=STREET))
pp<-pp %>% mutate(STREET=ifelse(STREET=="Pavd",yes="Paved",no=STREET))
pp$STREET<-as.factor(pp$STREET)
```

##MZZONE variable

```{r, message=FALSE, warning=FALSE}
library(sqldf)
table(pp$MZZONE)
SPMZ<-sqldf("select MZZONE,avg(SALES_PRICE) as SP  from pp group by MZZONE ")
SPMZ_br<-ggplot(data=SPMZ,aes(x=MZZONE,y=SP/1000))+geom_bar(fill="darkblue",stat = "Identity",width = 0.5)
SPISMZ<-ggplot(data=pp,aes(x=INT_SQFT,y=SALES_PRICE/1000,col=MZZONE))+geom_jitter(size=1)
grob5<-list(SPMZ_br,SPISMZ)
marrangeGrob(grob5,nrow = 1,ncol =1)

```


##Quality variables

```{r}
par(mfrow=c(2,2))
hist(pp$QS_ROOMS,col = "red")
hist(pp$QS_BATHROOM,col = "blue")
hist(pp$QS_BEDROOM,col = "red")
hist(pp$QS_OVERALL,col = "blue")
```

##Sq.ft to number of rooms

```{r}
pp$avg_sqft<-pp$INT_SQFT/pp$N_ROOM
summary(pp$avg_sqft)
par(mfrow=c(1,2))
hist(pp$avg_sqft,col="blue")
hist(sqrt(pp$avg_sqft),col="red")

```


##Missing values in Overall quality variable

```{r}
summary(pp$QS_OVERALL)
pp[is.na(pp$QS_OVERALL),]
pp$avg_qly<-apply(X = pp[15:17],MARGIN = 1,mean)
cor(x =pp[,23],y = pp[,18],use = "complete.obs")
ggplot(data=pp,aes(x=avg_qly))+geom_histogram(fill="darkblue",bins = 50)
```

##Imputing the Overall quality variable using linear regression

```{r, message=FALSE, warning=FALSE, include=FALSE}
set.seed(290391)
library(mice)
imp_pp_qly<-mice(pp[,c(18,23)],method = "norm.boot",m = 10,maxit=10,seed = 290391)
imp_pp_qly<-complete(imp_pp_qly)
pp[,c(18)]<-imp_pp_qly$QS_OVERALL
```

##REG_FEE

```{r, warning=FALSE}
summary(pp$REG_FEE)
cor(x =pp[,21],y = pp[,19],use = "complete.obs")
par(mfrow=c(1,2))
hist(pp$REG_FEE,col="blue")
hist(sqrt(pp$REG_FEE),col="red")
ggplot(data=pp,aes(x=REG_FEE/1000,y=SALES_PRICE/1000))+geom_point(size=1,shape=15,color="blue")+geom_smooth(method=lm,se=FALSE,fullrange=TRUE)
```

##Creating a variable called Guideline value as Registration fee is 1% of it

```{r}
pp$Guide_val<-pp$REG_FEE/0.01
summary(pp$Guide_val)
ggplot(data=pp,aes(y=Guide_val/1000,x=total_yrs/365,col=INT_SQFT))+geom_jitter(size=1,shape=15)
```

##Commission paid_COMMIS

```{r, warning=FALSE}
summary(pp$COMMIS)
cor(x =pp[,21],y = pp[,20],use = "complete.obs")
par(mfrow=c(1,2))
hist(pp$COMMIS,col="blue")
hist(sqrt(pp$COMMIS),col="red")
ggplot(data=pp,aes(x=COMMIS/1000,y=SALES_PRICE/1000))+geom_point(size=1,shape=15,color="red")+geom_smooth(method=lm,se=FALSE,fullrange=TRUE)

```

##Stamp Duty

```{r}
pp$Stamp_duty<-pp$Guide_val*0.08
summary(pp$Stamp_duty)
ggplot(data=pp,aes(y=Guide_val/1000,x=total_yrs/365,col=INT_SQFT))+geom_jitter(size=1,shape=15)
```

##Other charges which is Reg_fee,stamp_duty and commission
```{r}
pp$oth_chgs<-apply(pp[,c(19,20,26)],MARGIN = 1,sum)
summary(pp$oth_chgs)
```

##Sales_Price

```{r}
summary(pp$SALES_PRICE)
par(mfrow=c(1,2))
hist(pp$SALES_PRICE,col="red")
hist(sqrt(pp$SALES_PRICE),col="blue")
```

##Age of the house _ total_yrs

```{r, warning=FALSE}
summary(pp$total_yrs)
par(mfrow=c(1,2))
hist(pp$total_yrs,col="blue")
hist(sqrt(pp$total_yrs),col="red")

YRSP<-ggplot(data=pp,aes(x=total_yrs/365,y=SALES_PRICE/1000))+geom_point(size=1,shape=15,color="blue")+geom_smooth(method=lm,se=FALSE,color="darkred")
YRCO<-ggplot(data=pp,aes(x=total_yrs/365,y=COMMIS/1000))+geom_point(size=1,shape=15,color="red")+geom_smooth(method=lm,se=FALSE)
grob1<-list(YRSP,YRCO)
marrangeGrob(grob1,nrow = 1,ncol =2 )
```
##Removing the some variables

```{r}
library(caret)
#pp_cplot<-cplot(data = pp[,-21])
#findCorrelation(pp_cplot,cutoff = 0.6,names = TRUE)
pp_back<-pp
```

##

```{r}
pp<-pp_back
pp<-pp[,-c(3,10,15:17,19,24,25,26)]
```

##SQRT transformation of the sales price variable

```{r}
pp$SALES_PRICE<-sqrt(pp$SALES_PRICE)

pp$SALES_PRICE<-(pp$SALES_PRICE)^2
```

##Binning important variables

```{r}
bin_insq<-cut(pp$INT_SQFT,labels=1:4,breaks=c(0,1000,1500,2000,2500),ordered_result=FALSE)
pp$bin_insq<-bin_insq

##Binning distance from the main road

bin_dist<-cut(pp$DIST_MAINROAD,labels=1:4,breaks=c(-1,50,100,150,200),ordered_result=FALSE)
pp$bin_dist<-bin_dist

##Binning total yrs

bin_ttl_yrs<-cut(pp$total_yrs/365,labels=1:3,breaks=c(0,20,40,55),ordered_result=FALSE)
pp$bin_ttl_yrs<-bin_ttl_yrs

##Binning commission

bin_comis<-cut(pp$COMMIS/1000,labels=1:3,breaks=c(0,75,150,525),ordered_result=FALSE)

```

##Collapsing MZZONE variable

```{r, warning=FALSE}
#MZ_temp<-pp %>% mutate(MZZONE=ifelse(MZZONE=="A"|MZZONE=="C"|MZZONE=="I",yes="others",no=MZZONE))
#pp$AREA<-as.character(pp$AREA)
##pp<-pp %>% mutate(AREA=ifelse(AREA=="Adyar"|AREA=="Chrompet",yes="ADC",no=AREA))
##pp<-pp%>% mutate(AREA=ifelse(AREA=="Anna Nagar"|AREA=="T Nagar",yes="ANT",no=AREA))
#AR_temp$AREA<-as.factor(AR_temp$AREA)
#pp$AREA<-as.factor(pp$AREA)
```

##Interaction of the categorical variables

```{r, warning=FALSE}
pp$ARBT<-interaction(pp$AREA,pp$BUILDTYPE)
pp$INBT<-interaction(pp$BUILDTYPE,pp$bin_insq)
pp$TRBT<-interaction(pp$BUILDTYPE,pp$N_ROOM)
pp$BRBT<-interaction(pp$BUILDTYPE,pp$N_BEDROOM)
pp$ARBR<-interaction(pp$AREA,pp$N_ROOM)
pp$CMBR<-interaction(bin_comis,pp$N_ROOM)
```

##Some changes to BR_BHR variable because of unwanted levels

```{r, warning=FALSE}
dmy_pp<-dummyVars(data=pp,formula =~ARBR,levelsOnly = TRUE,fullRank = TRUE)
pp_dmy_pre<-predict(dmy_pp,pp)
pp<-cbind(pp[,-26],pp_dmy_pre)
dmy_rm<-nearZeroVar(x = pp,freqCut = 100/0,names = FALSE)
pp<-pp[,-c(dmy_rm)]

dmy_pp<-dummyVars(data=pp,formula =~CMBR,levelsOnly = TRUE,fullRank = TRUE)
pp_dmy_pre<-predict(dmy_pp,pp)
pp<-cbind(pp[,-26],pp_dmy_pre)
dmy_rm<-nearZeroVar(x = pp,freqCut = 100/0,names = FALSE)
pp<-pp[,-c(dmy_rm)]

```

##Bringing the train and test datasets

```{r}
pp_train<-pp[!is.na(pp$SALES_PRICE),]
pp_test<-pp[is.na(pp$SALES_PRICE),]

```

##Model
```{r}
set.seed(290391)

tr_cntl=trainControl(method="cv",number=5)
lm_pp<- train(SALES_PRICE~.,data=pp_train,
          method = "ridge",trControl=tr_cntl,preProcess=c("scale","center"))
lm_pp_pred<-predict(object = lm_pp,newdata = pp_test)
lm_pp
```

##Submission

```{r}
#submitlasso <- data.frame(PRT_ID=pp_temp1$PRT_ID,SALES_PRICE=lm_pp_pred)
#write.csv(submitlasso, "submitlasso.csv",row.names = FALSE)

```

##XGBOOST
```{r, eval=FALSE, include=FALSE}
# set.seed(290391)
# pp_train1_fea<-pp_train[,c(1,2,3,4,5,6,7,8,9,11,13,14,16:26,15)]
# pp_test_fea<-pp_test[,c(1,2,3,4,5,6,7,8,9,11,13,14,16:26,15)]
# 
# dmy_pp<-dummyVars(data=pp_train1_fea[,-24],formula =~.,levelsOnly = FALSE,fullRank = TRUE)
# pp_dmy_pre<-predict(dmy_pp,pp_train1_fea)
# pp_train2<-cbind(pp_train1_fea[,25],pp_dmy_pre)
# 
# dmy_pp<-dummyVars(data=pp_test_fea[,-24],formula =~.,levelsOnly = FALSE,fullRank = TRUE)
# pp_dmy_pre<-predict(dmy_pp,pp_test_fea)
# pp_test2<-cbind(pp_test_fea[,24],pp_dmy_pre)
# 
# 
# ControlParamteres <- trainControl(method = "cv",
#                   number = 5)
# 
# parametersGrid <-  expand.grid(eta=c(0.2,0.5),nrounds=100,lambda=c(0.5,0.9),alpha=c(0.5,0.9))
# 
# xgb_pp <- caret::train(y=pp_train2[,1],
#               x=pp_train2[,-c(1)],method="xgbLinear",
#                              trControl=ControlParamteres,tuneGrid=parametersGrid)
# 
# 
# xgb_pp_pred<- predict(xgb_pp, newdata=pp_test2)
# xgb_pp
```
```{r}
index=sample(x = nrow(pp_train),size = 5000)
X_train=pp_train[index,]
y_test=pp_train[-index,]

pred_train=predict(lm_pp,newdata = X_train[,-15])

RMSE(pred_train,obs = X_train$SALES_PRICE)

pred_test=predict(lm_pp,newdata = y_test[,-15])

RMSE(pred_test,obs = y_test$SALES_PRICE)


```
```{r}
library(devtools)
#install_github("h2oai/h2o-3/h2o-r/ensemble/h2oEnsemble-package")
require(h2oEnsemble)
# h2o.init(nthreads = -1)
# h2o.removeAll()
# 
# write.csv(X_train,file = 'Xtra.csv')
# write.csv(y_test,file = 'ytest.csv')
# tr1=h2o.importFile(path = 'Xtra.csv')
# tes1=h2o.importFile(path = 'ytest.csv')


```

```{r}
# learner <- c("h2o.randomForest.wrapper", 
#              "h2o.gbm.wrapper")
# metaLearner=c("h2o.gbm.wrapper")

#y='SALES_PRICE'


#setdiff(names(tr1),y)


# fit <- h2o.ensemble(x = 1:(ncol(tr1)-1), y = 'C57', 
#                     training_frame = tr1, 
#                     family = 'gaussian', 
#                     learner = learner, 
#                     metalearner = metaLearner,
#                     cvControl = list(V = 3))
# 
# fit <- h2o.ensemble(x = 1:(ncol(train_frame.hex)-1), y = ncol(train_frame.hex), 
#                     train_frame.hex, validation_frame=valid_frame.hex,
#                     family = "gaussian", 
#                     learner = learner, 
#                     metalearner = metalearner,
#                     cvControl = list(V = 5))	     
# 


```

```{r}
# pred=predict(object = fit,newdata = tes1)
# pred$pred
# as.data.frame(pred$pred) 

# head(pred)
```

